---
title: "Extended Mapping Assignment"
author: "Fan Feng"
date: "2020/11/11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
pacman::p_load("tidyverse",
               "magrittr",
               "hurricaneexposuredata",
               "drat",
               "maps",
               "lubridate",
               "kableExtra")
```

# 1. Background 

Main Task:

The purpose of this assignment is to make sure that you know the basics of using R to produce documents, presentations, and shiny apps. These are basic skills that every data scientists must have. Use the mapping project you have just completed as a starting point. I have provided additional county level data from FEMA and FEC so that you can produce maps with more features.

Using this data, produce a document using rmarkdown, a presentation using revealjs, and a Shiny application that you publish on shinyapps.io. Your document should include structure elements such as headings, lists, tables. Your presentation should also include these elements. Your Shiny application should allow users to select of adjust the display of maps and data plots. All of these deliverables should contain maps and data plots. You don't need to go overboard. Do enough to demonstrate your skill.

# 2. Data Cleaning

## Import the dataset & Select columns
```{r}
fema = read.csv("clean_data.csv")
```

```{r}
fema$state <- tolower(fema$state)
fema$county <- tolower(fema$county)
fema %<>% select(disasterNumber,date,damageCategory,projectSize,
                 state,county, projectAmount,federalShareObligated)
names(fema) <- c("hurricane_id","date","damage_category","project_size",
                 "region", "subregion", "project_amount", "federal_share")
```

## Name each hurricane

Based on the information from the FEMA website, we 
```{r}
#year18<-fema %>% filter(year(date)==2018)
#year18$hurricane_id %>% unique()
#2009
IDA <- fema %>% filter(hurricane_id == 1866) %>%
  mutate(hurr_name = 'IDA')
#2010
EARL <- fema %>% filter(hurricane_id %in% c(1939,3314,3315))%>%
  mutate(hurr_name = 'EARL')
ALEX <- fema %>% filter(hurricane_id == 1931) %>%
  mutate(hurr_name = 'ALEX')
#2011
IRENE <- fema %>% filter(year(date) == 2011) %>% 
  mutate(hurr_name = 'IRENE')
#2012-2013
SANDY <- fema %>% filter((year(date) == 2012) | (year(date) == 2013)) %>%
  mutate(hurr_name = 'SANDY')
#2016
MATTHEW <- fema %>% 
  filter(hurricane_id %in% c(4291,4286,4285,4283,4284)) %>%
  mutate(hurr_name = 'MATTHEW')
HERMINE <- fema %>% filter(hurricane_id == 4280) %>% 
  mutate(hurr_name = 'HERMINE')
#2017
HARVEY <- fema %>% filter(hurricane_id %in% c(4332,4345)) %>% 
  mutate(hurr_name = 'HARVEY')
NATE <- fema %>% filter(hurricane_id %in% c(3392,3395,4349,4350)) %>% 
  mutate(hurr_name = 'NATE')
IRMA <- fema %>%  
  filter(hurricane_id %in% c(3384,3389,4336,4337,4338,4346,4335,4341)) %>%
  mutate(hurr_name = 'IRMA')
MARIA <- fema %>% filter(hurricane_id %in% c(4340,4339)) %>% 
  mutate(hurr_name = 'MARIA')
#2018
MICHAEL <- fema %>% filter(hurricane_id %in% c(4411,4406,4400,4399)) %>%
  mutate(hurr_name = 'MICHAEL')
FLORENCE <- fema %>% filter(hurricane_id %in% c(4393,4394,4401)) %>%
  mutate(hurr_name = 'FLORENCE')
LANE_hawaii <- fema %>% filter(hurricane_id == 4395) %>%
  mutate(hurr_name = 'LANE')
GITA <- fema %>% filter(hurricane_id == 4357) %>%
  mutate(hurr_name = 'GITA')
```

## Combine all the hurricanes 
```{r}
fema_new <- do.call("rbind", list(IDA,EARL,ALEX,IRENE,SANDY,MATTHEW,
                                  HERMINE,HARVEY,NATE,IRMA,MARIA,MICHAEL,
                                  FLORENCE,LANE_hawaii,GITA))
#write.csv(fema_new, "Shiny/shiny_data.csv")
```

# 3. Exploratory Data Analysis

## 3.1 Temporal Aspect

First, we explore our data from the aspect of time. We draw multiple plots based on the data of different years to reveal various information.

### The number of major disaster declaration each year
```{r, echo = FALSE}
year_count <- fema_new %>%
  mutate(Year = year(date)) %>% 
  group_by(Year) %>%
  tally()
```

```{r, echo = FALSE}
ggplot(year_count,
       aes(x = as.factor(Year), y = n, fill = as.factor(Year))) +
  geom_bar(stat = 'identity') +
  scale_fill_hue(c = 40) +
  ggtitle("The Number of Major Hurricane Delcaration Each Year") +
  xlab("Year") + ylab("The Count of Delcaration") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5))
```

```{r,echo=FALSE}
ggplot(data = year_count,
       mapping = aes(x = as.factor(Year), y = n, group = 1)) + 
  geom_line() + geom_point() + 
  scale_fill_hue(c = 40) +
  ggtitle("The Change of Declaration's Count over Years") +
  xlab("Year") + ylab("The Count of Hurricane Declarations") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
```

### The Number of States Suffered From Hurricane Each Year
```{r,echo=FALSE}
year_state <- fema_new %>%
  mutate(Year = year(date)) %>%
  group_by(Year,region) %>%
  tally()
```

```{r,echo=FALSE}
ggplot(year_state) +
  geom_bar(aes(x = as.factor(Year),fill = as.factor(Year))) +
  scale_fill_hue(c = 40) +
  ggtitle("The Number of States Suffered From Hurricane Each Year") +
  xlab("Year") + ylab("The Count of States Suffered") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
  #coord_flip()
```

### The Total Project Amount Each Year
```{r, ,echo=FALSE}
year_amount <- fema_new %>% 
  mutate(Year = year(date)) %>%
  group_by(Year) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

severe_year = year_amount[rev(order(year_amount$sum))[1:5],]
peace_year = year_amount[rev(order(year_amount$sum))[6:8],]
```

```{r,echo=FALSE}
ggplot(data = year_amount,
       mapping = aes(x = as.factor(Year), y = sum, 
                     fill = as.factor(Year))) + 
  geom_bar(stat = 'identity') + 
  scale_fill_hue(c = 40) +
  ggtitle("The Total Project Amount Each Year") +
  xlab("Year") + ylab("The Total Project Amount") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
```

```{r,echo=FALSE}
ggplot(data = severe_year,
       mapping = aes(x = as.factor(Year), y = sum, 
                     fill = as.factor(Year))) + 
  geom_bar(stat = 'identity') + 
  scale_fill_hue(c = 40) +
  ggtitle("The Total Project Amount - Severe Years") +
  xlab("Year") + ylab("The Total Project Amount") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
```

```{r,echo=FALSE}
ggplot(data = peace_year,
       mapping = aes(x = as.factor(Year), y = sum, 
                     fill = as.factor(Year))) + 
  geom_bar(stat = 'identity') + 
  scale_fill_hue(c = 40) +
  ggtitle("The Total Project Amount - Relatively Peaceful Years") +
  xlab("Year") + ylab("The Total Project Amount") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
```

```{r,echo=FALSE}
ggplot(data = year_amount,
       mapping = aes(x = as.factor(Year), y = sum, group = 1)) + 
  geom_line() + geom_point() + 
  scale_fill_hue(c = 40) +
  ggtitle("The Change of Total Project Amount over Years") +
  xlab("Year") + ylab("The Total Project Amount") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
```

## 3.2 Spatial Aspect

### The Total Project Amount Each State ALL These Years
```{r,echo=FALSE}
state_level <- fema_new %>% filter(subregion=="statewide")
state_amount <- state_level %>% 
  group_by(region) %>% 
  summarize(sum = sum(project_amount),.groups='drop')
```

```{r,echo=FALSE}
ggplot(data = state_amount, 
       aes(x = reorder(region,sum), y = sum, fill = region)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  ylab("The Total Project Amount") + 
  ggtitle("Total Project Amount Each State ALL These Years") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```

```{r,echo=FALSE}
severe_state <- state_amount[rev(order(state_amount$sum))[1:5],]
ggplot(data = severe_state, 
       aes(x = reorder(region,sum), y = sum, fill = region)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  ylab("The Total Project Amount") + 
  ggtitle("Top 5 Total Project Amount Each State ALL These Years") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```

### The Total Project Amount Each County ALL These Years
```{r,warning=FALSE,echo=FALSE}
county_level <- fema_new %>% filter(subregion!= "statewide")
county_amount <- county_level %>% 
  group_by(subregion) %>% 
  summarize(sum = sum(project_amount),.groups='drop')
```

```{r,echo=FALSE}
ggplot(data = county_amount[rev(order(county_amount$sum))[1:5],], 
       aes(x = reorder(subregion,sum), y = sum, fill = subregion)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  ylab("The Total Project Amount") + 
  ggtitle("Top 10 Total Project Amount Each County ALL These Years") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```

## 3.3 Hurriance Aspect

### The Total Project Amount of Each Hurricane 
```{r,warning=FALSE,echo=FALSE}
hurr_level <- fema_new %>% filter(subregion != "statewide")
hurr_amount <- hurr_level %>% 
  group_by(hurr_name) %>% 
  summarize(sum = sum(project_amount),.groups='drop')
```

```{r,echo=FALSE}
ggplot(data = hurr_amount, 
       aes(x = reorder(hurr_name,sum), y = sum, fill = hurr_name)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  ylab("The Total Project Amount") + 
  ggtitle("The Total Project Amount of Each Hurricane") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```

```{r,echo=FALSE}
ggplot(data = hurr_amount[rev(order(hurr_amount$sum))[1:5],], 
       aes(x = reorder(hurr_name,sum), y = sum, fill = hurr_name)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  ylab("The Total Project Amount") + 
  ggtitle("Top 10 Total Project Amount of Each Hurricane") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```

## 3.4 project size aspect

We have two categories of project size: Large and Small.
```{r}
size_amount <- fema_new %>% 
  group_by(project_size) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

size_count <- fema_new %>% 
  group_by(project_size) %>% 
  summarize(count = n(),.groups='drop')
```

### The Total Project Amount for Each Project Size
```{r}
ggplot(data = size_amount, 
       aes(x = factor(project_size), y = sum, 
           fill = factor(project_size))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("Project Size") + 
  ylab("Sum of Project Amount") + 
  ggtitle("The Total Project Amount for Each Project Size") + 
  theme(plot.title = element_text(hjust = 0.5))
```

### The Frequencies for Each Project Size
```{r}
ggplot(data = size_count, 
       aes(x = factor(project_size), y = count, 
           fill = factor(project_size))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) + 
  xlab("project size") + 
  ylab("count of project size") + 
  ggtitle("The Frequencies for Each Project Size") + 
  theme(plot.title = element_text(hjust = 0.5))
```

## 3.5 Damage Category Aspect
```{r}
dcc_amount <- fema_new %>% 
  group_by(damage_category) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

dcc_count <- fema_new %>% 
  group_by(damage_category) %>% 
  summarize(count = n(),.groups='drop')
```

### Total Project Amount for Each Damage Category
```{r}
ggplot(data = dcc_amount, 
       aes(x = reorder(damage_category,sum), y = sum, 
           fill = as.factor(damage_category))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("Damage Catrgory") + 
  ylab("Total Project Amount") + 
  ggtitle("Total Project Amount for Each Damage Category") + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") +
  coord_flip()
```

### The Frequencies for Each Damage Category

```{r}
ggplot(data = dcc_count, 
       aes(x = reorder(damage_category,count), y = count, 
           fill = as.factor(damage_category))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("Damage Catrgory") + 
  ylab("Frequency") + 
  ggtitle("The Frequencies for Each Damage Category") + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") +
  coord_flip()
```

# 4. Ploting the maps

First, let’s check whether the dataset we used to make the maps contains all our regions in the fema dataset.
```{r}
all <- map_data("county")
a <- unique(all$region)
b <- unique(fema_new$region)
b[!(b %in% a)]
```

We notice that there are four regions not included in the map data.

Then, we load the data of longitude and latitude, and join them with our data
```{r}
sandy_map <- merge(all,SANDY,by=c("region","subregion"))
irma_map <- merge(all,IRMA,by=c("region","subregion"))
#maria_map <- merge(all,MARIA,by=c("region","subregion"))
harvey_map <- merge(all,HARVEY,by=c("region","subregion"))
michael_map <- merge(all,MICHAEL,by=c("region","subregion"))
```

### Make map plots of the project amount of Hurricane Sandy
```{r}
sandyPlot <- ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=sandy_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount (USD)") + 
  ggtitle("Sandy") +
  theme(plot.title = element_text(hjust = 0.5))
sandyPlot
```


### Make map plots of the project amount of Hurricane IRMA
```{r}
irmaPlot <- ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=irma_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount") + 
  ggtitle("IRMA") +
  theme(plot.title = element_text(hjust = 0.5))
irmaPlot
```

### Make map plots of the project amount of Hurricane Harvey
```{r}
harveyPlot <- ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=harvey_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount (USD)") + 
  ggtitle("Harvey") +
  theme(plot.title = element_text(hjust = 0.5))
harveyPlot
```

### Make map plots of the project amount of Hurricane Michael
```{r}
michaelPlot <- ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=michael_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount (USD)") + 
  ggtitle("Michael") +
  theme(plot.title = element_text(hjust = 0.5))
michaelPlot
```


# 5. Shiny App

[Click here go to shiny app](http://)

# 6. Acknowledgements $ References

The data is collected from: https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1

1. Yihui Xie (2020). knitr: A General-Purpose Package for Dynamic Report Generation in R. R package version 1.29. https://github.com/yihui/knitr

2. Hadley Wickham (2019). tidyverse: Easily Install and Load the 'Tidyverse'. R package version 1.3.0. https://cloud.r-project.org/package=tidyverse

3. Greg Freedman Ellis, Thomas Lumley, Tomasz Żółtak and Ben Schneider (2020). srvyr:'dplyr'-Like Syntax for Summary Statistics of SUrvey Data. R package version 0.4.0. https://cloud.r-project.org/package=srvyr


