---
title: "MA615 Hurricane Project"
author: "Fan Feng"
date: 2020/11/12
output: 
  revealjs::revealjs_presentation:
  theme: serif
  transition: slide
  self_contained: false
  reveal_plugins: ["menu"]
  center: false
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>",
                      collapse = TRUE,
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      dpi = 300,
                      cache.lazy = FALSE,
                      tidy = "styler",
                      out.width = "90%",
                      fig.align = "center",
                      fig.width = 10,
                      fig.height = 4)
pacman::p_load(usmap, dplyr, tidyverse, drat, hurricaneexposuredata, ggplot2, maps, tmap, leaflet, sf, sp, lubridate, readr, kableExtra, magrittr, gridExtra, knitr,revealjs,knitr,png,grid)
options(crayon.enabled = FALSE)
theme_set(theme_light())
```

# Introduction

The purpose of this assignment is to show that we know the basics of using R to produce documents, presentations, and shiny apps. These are basic skills that every data scientists must have. 

In this assignment, we use the mapping project you have just completed as a starting point. In addition, we have more county level data from FEMA and FEC so that we can produce maps with relevant features.

# Data Description

The Public Assistance Funded Projects Details dataset contains obligated (financial obligation to grantee) Public Assistance projects.

FEMA provides supplemental Federal disaster grant assistance for debris removal, emergency protective measures, and the repair, replacement, or restoration of disaster-damaged, publicly owned facilities and the facilities of certain Private Non-Profit (PNP) organizations through the PA Program. The PA Program also encourages protection of these damaged facilities from future events by providing assistance for hazard mitigation measures during the recovery process.

## Data Cleaning

-We download raw data [PublicAssistanceFundedProjectsDetails.csv]( https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1) from the website of FEMA and load into R. 

- Filter the data from 2009 ~ 2018
- Add the Names all the hurricanes
- Remove the redundant columns 

```{r, echo = FALSE}
fema_new = read.csv("slides_data.csv")
```

# Exploratory Data Analysis

- 1.Temporal Aspect
- 2.Spatial Aspect
- 3.Hurricane Aspect
- 4.Project Size Aspect
- 5.Damage Category Aspect

## 1. Temporal Aspect

### The number of major disaster declaration each year
<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
year_count <- fema_new %>%
  mutate(Year = year(date)) %>% 
  group_by(Year) %>%
  tally()

ggplot(year_count,
       aes(x = as.factor(Year), y = n, fill = as.factor(Year))) +
  geom_bar(stat = 'identity') +
  scale_fill_hue(c = 40) +
  ggtitle("The Number of Major Hurricane Delcaration Each Year") +
  xlab("Year") + ylab("The Count of Delcaration") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5))
```
</div>

## The Number of States Suffered From Hurricane Each Year

<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
year_state <- fema_new %>%
  mutate(Year = year(date)) %>%
  group_by(Year,region) %>%
  tally()
ggplot(year_state) +
  geom_bar(aes(x = as.factor(Year),fill = as.factor(Year))) +
  scale_fill_hue(c = 40) +
  ggtitle("The Number of States Suffered From Hurricane Each Year") +
  xlab("Year") + ylab("The Count of States Suffered") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
```
<div>

## The Total Project Amount Each Year
<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
year_amount <- fema_new %>% 
  mutate(Year = year(date)) %>%
  group_by(Year) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

severe_year = year_amount[rev(order(year_amount$sum))[1:5],]
peace_year = year_amount[rev(order(year_amount$sum))[6:8],]

ggplot(data = year_amount,
       mapping = aes(x = as.factor(Year), y = sum, 
                     fill = as.factor(Year))) + 
  geom_bar(stat = 'identity') + 
  scale_fill_hue(c = 40) +
  ggtitle("The Total Project Amount Each Year") +
  xlab("Year") + ylab("The Total Project Amount") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 
```
<div>

## Severe Years & Peaceful Years
<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
p1 <- ggplot(data = severe_year,
             mapping = aes(x = as.factor(Year), y = sum, 
                           fill = as.factor(Year))) + 
  geom_bar(stat = 'identity') + 
  scale_fill_hue(c = 40) +
  ggtitle("The Total Project Amount - Severe Years") +
  xlab("Year") + ylab("The Total Project Amount") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5)) 

p2 <- ggplot(data = peace_year,
             mapping = aes(x = as.factor(Year), y = sum, 
                          fill = as.factor(Year))) + 
  geom_bar(stat = 'identity') + 
  scale_fill_hue(c = 40) +
  ggtitle("The Total Project Amount - Relatively Peaceful Years") +
  xlab("Year") + ylab("The Total Project Amount") +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5))

grid.arrange(p1,p2,nrow=1)
```
<div>

## 2. Spatial Aspect

### The Total Project Amount Each State ALL These Years
<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
state_level <- fema_new %>% filter(subregion=="statewide")
state_amount <- state_level %>% 
  group_by(region) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

ggplot(data = state_amount, 
       aes(x = reorder(region,sum), y = sum, fill = region)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("State") +
  ylab("The Total Project Amount") + 
  ggtitle("Total Project Amount Each State ALL These Years") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```
<div>

## Top 5 States & Top 5 Counties
<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
severe_state <- state_amount[rev(order(state_amount$sum))[1:5],]
p3 <- ggplot(data = severe_state, 
             aes(x = reorder(region,sum), y = sum, fill = region)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("State") +
  ylab("The Total Project Amount") + 
  ggtitle("Top 5 States in Total Project Amount") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```
<div>

```{r,warning=FALSE,echo=FALSE}
county_level <- fema_new %>% filter(subregion!= "statewide")
county_amount <- county_level %>% 
  group_by(subregion) %>% 
  summarize(sum = sum(project_amount),.groups='drop')
```

<div style="font-size:10pt; font-weight: bold">
```{r,echo=FALSE}
p4 <- ggplot(data = county_amount[rev(order(county_amount$sum))[1:5],], 
             aes(x = reorder(subregion,sum), y = sum, fill = subregion)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("County") +
  ylab("The Total Project Amount") + 
  ggtitle("Top 5 County in Total Project Amount ") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
grid.arrange(p3,p4,nrow=1)
```
<div>

## 3. Hurriance Aspect

### The Total Project Amount of Each Hurricane 
<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
hurr_level <- fema_new %>% filter(subregion != "statewide")
hurr_amount <- hurr_level %>% 
  group_by(hurr_name) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

ggplot(data = hurr_amount, 
       aes(x = reorder(hurr_name,sum), y = sum, fill = hurr_name)) +
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("Hurricane Name") +
  ylab("The Total Project Amount") + 
  ggtitle("The Total Project Amount of Each Hurricane") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")+
  coord_flip()
```
<div>

## 4. Project Size Aspect

<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
size_amount <- fema_new %>% 
  group_by(project_size) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

size_count <- fema_new %>% 
  group_by(project_size) %>% 
  summarize(count = n(),.groups='drop')

p5<- ggplot(data = size_amount, 
            aes(x = factor(project_size), y = sum, fill = factor(project_size))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("Project Size") + 
  ylab("Sum of Project Amount") + 
  ggtitle("The Total Project Amount for Each Project Size") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")

p6<-ggplot(data = size_count, 
           aes(x = factor(project_size), y = count, 
               fill = factor(project_size))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) + 
  xlab("project size") + 
  ylab("count of project size") + 
  ggtitle("The Frequencies for Each Project Size") + 
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none") 
grid.arrange(p5,p6,nrow=1)
```
<div>

## 5. Damage Category Aspect

<div style="font-size:10pt; font-weight: bold">
```{r, echo = FALSE}
dcc_amount <- fema_new %>% 
  group_by(damage_category) %>% 
  summarize(sum = sum(project_amount),.groups='drop')

dcc_count <- fema_new %>% 
  group_by(damage_category) %>% 
  summarize(count = n(),.groups='drop')

p7<- ggplot(data = dcc_amount, 
            aes(x = reorder(damage_category,sum), y = sum, 
                fill = as.factor(damage_category))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("Damage Catrgory") + 
  ylab("Total Project Amount") + 
  ggtitle("Total Project Amount for Each Damage Category") + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") +
  coord_flip()

p8 <- ggplot(data = dcc_count, 
             aes(x = reorder(damage_category,count), y = count, 
                 fill = as.factor(damage_category))) + 
  geom_bar(stat = "identity") + 
  scale_fill_hue(c = 40) +
  xlab("Damage Catrgory") + 
  ylab("Frequency") + 
  ggtitle("The Frequencies for Each Damage Category") + 
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") +
  coord_flip()
grid.arrange(p7,p8,nrow=1)
```
<div>

# Mapping The Hurricanes

1. Combine fema_new dataset with map data.

2. Choose Top 4 destructive hurricanes to plot maps

## Maps of Hurricanes
<div style="font-size:18pt">
<p>We draw the total cost of most destructive hurricanes in map.</p>  
```{r echo=FALSE, message=FALSE}

all <- map_data("county")
SANDY <- fema_new %>% filter(hurr_name == 'SANDY')
IRMA <- fema_new %>% filter(hurr_name == 'IRMA')
HARVEY <- fema_new %>% filter(hurr_name == 'HARVEY')
MICHAEL <- fema_new %>% filter(hurr_name == 'MICHAEL')

sandy_map <- merge(all,SANDY,by=c("region","subregion"))
irma_map <- merge(all,IRMA,by=c("region","subregion"))
#maria_map <- merge(all,MARIA,by=c("region","subregion"))
harvey_map <- merge(all,HARVEY,by=c("region","subregion"))
michael_map <- merge(all,MICHAEL,by=c("region","subregion"))
```

## Hurricane - SANDY
```{r echo=FALSE, message=FALSE}
ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=sandy_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount (USD)") + 
  ggtitle("Sandy") +
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")
```

## Hurricane - IRMA
```{r echo=FALSE, message=FALSE}
ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=irma_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount") + 
  ggtitle("IRMA") +
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")

```

## Hurricane - Harvey
```{r echo=FALSE, message=FALSE}
ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=harvey_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount (USD)") + 
  ggtitle("Harvey") +
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")

```


## Hurricane - Michael
```{r echo=FALSE, message=FALSE}
ggplot() + 
  geom_polygon(data=all, 
               aes(x=long, y=lat, group=group),colour="black",fill="white")+
  geom_polygon(data=michael_map, 
               aes(x = long, y = lat, group = group, fill=project_amount))+
  labs(fill="Project Amount (USD)") + 
  ggtitle("Michael") +
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none")
```

# References

The data is collected from: https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1

1. Yihui Xie (2020). knitr: A General-Purpose Package for Dynamic Report Generation in R. R package version 1.29. https://github.com/yihui/knitr

2. Hadley Wickham (2019). tidyverse: Easily Install and Load the 'Tidyverse'. R package version 1.3.0. https://cloud.r-project.org/package=tidyverse

3. Greg Freedman Ellis, Thomas Lumley, Tomasz Żółtak and Ben Schneider (2020). srvyr:'dplyr'-Like Syntax for Summary Statistics of SUrvey Data. R package version 0.4.0. https://cloud.r-project.org/package=srvyr


# Thanks

Thanks for Listening!



